The Uneven-Height Two-Column Specht Modules of the Hecke Algebra of $S_n$

### 2019-07-18 tex/Generalized-Crossingless.tex
Hecke Algebra Relations Verifications

We will prove that $W := W_{2n+r}^r$ and $S := S^{(n+r,n)'}$ are isomorphic as representations in the case that $\SH$ is semisimple.

\newpage
\section{Correspondence} 
We can now begin by proving that $W$ is irreducible;
then $W \simeq S^\lambda$ for some partition $\lambda$ of $2n + r$, and we may use branching rules to determne $W$.
\begin{lemma}
  Every basis vector in $W_{2n + r}^r$ is cyclic.
\end{lemma}
\begin{proof}
  We have already proven this in the $r = 0$ case, so suppose that $r > 0$.

@@

\begin{proposition}
  \;

  \begin{enumerate}[label={(\roman*)}]
    \item The representation $W_{2+r}^r$ is reducible iff $e \mid r+2$.
    \item When $n \neq 1$ and $e > n+1$, the representation $W_{2n + r}^r$ is irreducible.
  \end{enumerate}
\end{proposition}
\begin{proof}
  \def\sectionsep{7pt}
  \textbf{(i)}
  Note that $\ima (1 + T_i)$ is 1-dimensional for each $i$, so it is equivalent that 
  \[K := \bigcap_{i=1}^{r+1} \ker(1 + T_i) = \ker \bigoplus_{i=1}^{r+1} (1 + T_i)\] 
  is trivial via the lemma.
  The transformation $\bigoplus (1 + T_i)$ is a linear operator on $W_{2 + r}^r$ given by the following matrix:
  \newcommand*\bigzero{\makebox(0,15){\text{\Huge0}}}
  \newcommand*\bigzerotwo{\makebox(-20,15){\text{\Huge0}}}
  \[
    A_{r + 1} = 
    \begin{bmatrix}
      \begin{matrix}
      q + 1 & q^{1/2} & \\
      q^{1/2} & q + 1 & q^{1/2}\\
      & q^{1/2} & q + 1 & q^{1/2}\\
      &  & \ddots & \ddots
      \end{matrix}
      &   \bigzero\\
      \bigzerotwo & \begin{matrix}
        q^{1/2} & q+1 & q^{1/2}\\
        &  q^{1/2} & q + 1
       \end{matrix}
    \end{bmatrix}
  \].
  Hence $K$ is trivial iff the determinant $\det A_{r + 1} = \brk{r + 2}_q$ is 0, or equivalently iff $e \mid r + 2$.

  \iffalse % Previous proof, kept for archival purposes
  \vspace{\sectionsep}
  \textbf{(ii)}
  We have already proven this for $r = 0$ and $n = 0$, so suppose $r > 0$ and $n > 1$.
  We will break into case work on $r$;
  suppose first that $r = 1$.
  Then, it is easy to verify that $W_{2n + 1}^1 \simeq \Res W_{2n + 1}^0$, which we have already proven irreducible.
  We may henceforth assume that $r > 1$.

@@

  Define the element
  \[
    h = (1 + T_{n + r - 1})(1 + T_{n+r-3})(1 + T_{n+r-2})(1 + T_{n + r - 4})\dots(1 + T_{a_{r-1} +1})(1 + T_{a_{r-1}+2}).
  \]
  Note that $h$ carries a basis element to a multiples of another basis element, and a basis element is in $\ker \, h$ iff it is not in $U$.
  Hence $hw'$ is a nonzero multiple of a basis element, giving $w$ cyclic.
  \fi

@@

  \begin{figure}
    \[
      \GeneralizedNAction{6}{1/2,4/5}{2}{3/1, 6/2}{3/4}{3/1, 2/2}{1/4, 2/3}{5/1, 6/2}{q}
      \hspace{50pt}
      \GeneralizedNAction{6}{1/2,4/5}{2}{3/1, 6/2}{3/4}{4/1, 3/2}{1/2, 3/4}{5/1, 6/2}{x}
    \]
  \caption{
    Demonstration of how the transformation $(1 + T_{i+1})(1 + T_i)$ ``moves'' anchors from positions $i,i+1$ to position $i+3$, with constant $x = q^{1/2}(q+1)$.
    Iterating this across all elements between $r+1$ and $2n + r$ via $h$ concentrates all anchors at the beginning or end; 
    if there are at least two anchors in $w_j$ after $r+1$, then we must act on two anchors eventually, giving $w_j \in \ker \, h$.
  }\label{h}
\end{figure}
\end{proof}

@@

\begin{proposition}
  Suppose that $n,r > 0$.
  Then, a filtration of $\Res W_{2n + r}^r$ is given by
  \begin{equation}
    0 \subset W_{2n + r - 1}^{r-1} \subset \Res \, W_{2n + r}^r \label{Filtration}
  \end{equation}
  with $\Res \, W_{2n + r}^r / W_{2n + r - 1}^{r-1} \simeq W_{2n + r - 1}^{r + 1}$. 
\end{proposition}
\begin{proof}
  Note that we may identify the subrepresentation of $\Res \; W_{2n + r}^r$ having anchor $n$ with $W_{2n + r - 1}^{r-1}$.

@@

\begin{corollary}
  When $\SH$ is semisimple, $W_{2n + r}^r \simeq S^{(n + r,r)'}$.
\end{corollary}
\begin{proof}
  We may argue by induction on $2n + r$, knowing that we have proven the base case $2n + r = 2$.
  Assume that we have proven the isomorphism for all $W_{2n + s}^{s}$ with $2n + s = 2n + r - 2$.
  We have proven the $n = 0$ and $r = 0$ cases already, so assume $n,r > 0$.

  Then, $W_{2n + r}^{r}$ is the unique irreducible representation of $\SH$ having restriction $S^{(n+r-1,n)'} \oplus S^{(n + r,n-1)'}$, implying the desired isomorphism.
\end{proof}

>>>>

We will prove that $W := W_{2n+r}^r$ and $S := S^{(n+r,n)'}$ are isomorphic as representations in the case that $e > n + r + 1$ is semisimple.

\newpage 
\section{Correspondence} 
\begin{proposition}
  Suppose that $n,r > 0$.
  Then, a filtration of $\Res W_{2n + r}^r$ is given by
  \begin{equation}
    0 \subset W_{2n + r - 1}^{r-1} \subset \Res \, W_{2n + r}^r \label{Filtration}
  \end{equation}
  with $\Res \, W_{2n + r}^r / W_{2n + r - 1}^{r-1} \simeq W_{2n + r - 1}^{r + 1}$. 
\end{proposition}
\begin{proof}
  Note that we may identify the subrepresentation of $\Res \; W_{2n + r}^r$ having anchor $n$ with $W_{2n + r - 1}^{r-1}$.

\begin{lemma}
  Every basis vector in $W_{2n + r}^r$ is cyclic.
\end{lemma}
\begin{proof}
  We have already proven this in the $r = 0$ case, so suppose that $r > 0$.

@@

\begin{proposition}
  The representation $W_{2n + r}^r$ is irreducible when $e > n + r + 1$. 
\end{proposition}
\begin{proof}
  \def\sectionsep{7pt}
  \iffalse % now unnecessary, kept for archival purposes.
  The representation $W_{2+r}^r$ is reducible iff $e \mid r+2$.
  \textbf{(i)}
  Note that $\ima (1 + T_i)$ is 1-dimensional for each $i$, so it is equivalent that 
  \[K := \bigcap_{i=1}^{r+1} \ker(1 + T_i) = \ker \bigoplus_{i=1}^{r+1} (1 + T_i)\] 
  is trivial via the lemma.
  The transformation $\bigoplus (1 + T_i)$ is a linear operator on $W_{2 + r}^r$ given by the following matrix:
  \newcommand*\bigzero{\makebox(0,15){\text{\Huge0}}}
  \newcommand*\bigzerotwo{\makebox(-20,15){\text{\Huge0}}}
  \[
    A_{r + 1} = 
    \begin{bmatrix}
      \begin{matrix}
      q + 1 & q^{1/2} & \\
      q^{1/2} & q + 1 & q^{1/2}\\
      & q^{1/2} & q + 1 & q^{1/2}\\
      &  & \ddots & \ddots
      \end{matrix}
      &   \bigzero\\
      \bigzerotwo & \begin{matrix}
        q^{1/2} & q+1 & q^{1/2}\\
        &  q^{1/2} & q + 1
       \end{matrix}
    \end{bmatrix}
  \].
  Hence $K$ is trivial iff the determinant $\det A_{r + 1} = \brk{r + 2}_q$ is 0, or equivalently iff $e \mid r + 2$.

  \vspace{\sectionsep}
  \textbf{(ii)}
  We have already proven this for $r = 0$ and $n = 0$, so suppose $r > 0$ and $n > 1$.
  We will break into case work on $r$;
  suppose first that $r = 1$.
  Then, it is easy to verify that $W_{2n + 1}^1 \simeq \Res W_{2n + 1}^0$, which we have already proven irreducible.
  We may henceforth assume that $r > 1$.

@@

  Define the element
  \[
    h = (1 + T_{n + r - 1})(1 + T_{n+r-3})(1 + T_{n+r-2})(1 + T_{n + r - 4})\dots(1 + T_{a_{r-1} +1})(1 + T_{a_{r-1}+2}).
  \]
  Note that $h$ carries a basis element to a multiples of another basis element, and a basis element is in $\ker \, h$ iff it is not in $U$.
  Hence $hw'$ is a nonzero multiple of a basis element, giving $w$ cyclic.

@@

  \begin{figure}
    \[
      \GeneralizedNAction{6}{1/2,4/5}{2}{3/1, 6/2}{3/4}{3/1, 2/2}{1/4, 2/3}{5/1, 6/2}{q}
      \hspace{50pt}
      \GeneralizedNAction{6}{1/2,4/5}{2}{3/1, 6/2}{3/4}{4/1, 3/2}{1/2, 3/4}{5/1, 6/2}{x}
    \]
  \caption{
    Demonstration of how the transformation $(1 + T_{i+1})(1 + T_i)$ ``moves'' anchors from positions $i,i+1$ to position $i+3$, with constant $x = q^{1/2}(q+1)$.
    Iterating this across all elements between $r+1$ and $2n + r$ via $h$ concentrates all anchors at the beginning or end; 
    if there are at least two anchors in $w_j$ after $r+1$, then we must act on two anchors eventually, giving $w_j \in \ker \, h$.
  }\label{h}
\end{figure}
  \fi
  We proceed by induction on $2n + r$.
  Note that, by identification with the trivial and sign representations, the base case $2n + r = 2$ is already prove, so suppose we have proven this for each $2m + s < 2n + r$.

  We will prove this in essentially the same way as before;
  the inductive argument continues to apply assuming that $K := \ker \, \bigoplus(1 + T_i)$ is trivial, and we will prove that this is trivial when $e \not| n + r + 1$ using a similar argument.
  This argument is very long, and I will recreate it later.
\end{proof}

@@

@@

\begin{proposition}
  Suppose $e > n + r + 1$, and let $\lambda$ be a partition of $2n + r$.
  \begin{enumerate}[label={(\roman*)}]
    \item Suppose $n,r > 0$.
      If $\Res \, S^\lambda$ admits a filtration $0 \subset S^{(n+r-1,n)'} \subset \Res \, S^\lambda$ with $\Res \, S^\lambda / S^{(n + r - 1,n)'} \simeq S^{(n + r,n-1)'}$ then $\lambda = (n+r,n)'$.
    \item Suppose $r = 0$.
      If $\Res \, S^\lambda \simeq S^{(n,n-1)'}$, then $\lambda = (n,n)'$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  \textbf{(i)}
  Recall that the characteristic-free classical branching theorem gives that every specht module $S^{\lambda}$ admits a filtration 
  \begin{equation}
    0 = M_0 \subset \dots \subset M_l = S^{\lambda} \label{filtration}
  \end{equation}  
    with $M_i/M_{i-1} \simeq S^{\lambda(i)}$.
  Hence $\lambda = (n + r, n)'$ satisfies the above formula.

  Note that, when $e > n + r + 1$, this filtration gives a composition series.
  Hence the sequence \eqref{filtration} is at most length-2.

  Suppose it is length-1.
  Then $\lambda$ is rectangle-shaped;
  to have rows removed to the above we then need that that it be $\lambda = \prn{\frac{2n + r}{2}, \frac{2n + r}{2}}'$, so that $\Res S^{\lambda} \simeq S^{\prn{\frac{2n + r}{2}, \frac{2n + r}{2} - 1}'}$.
  However, since $e > n + r + 1 \geq n + \frac{r}{2} + 1$, $\Res \, S^{\lambda}$ is irreducible, contradicting the existence of a length-2 composition series in the hypothesis.

  Now suppose that \eqref{filtration} is length 2, so that it is also a composition series.
  Then, by the Jordan-H\"older theorem, there is a rearranement of the composition factors of \ref{filtration} to give $S^{(n + r - 1,n)'}$ and $S^{(n + r, n - 1)'}$.
  This implies that both $(n + r - 1,n)'$ and $(n + r,n-1)'$ can be acquired by removing rows of $\lambda$, giving that $\lambda = (n+r,n)'$.

  \vspace{7pt}
  \textbf{(ii)}
  Note that, since $e > n + 1$, $\Res S^\lambda$ is irreducible.
  Hence the sequence \eqref{filtration} is a composition series of length 1, and $\lambda(1) = (n,n-1)'$.
  The first implies that $\lambda$ is rectangular, and the second implies that $\lambda = (n,n)'$.
\end{proof}

\begin{corollary}
  If $e > n + r + 1$, then $W_{2n + r}^r \simeq S^{(n + r,n)'}$.\qed
\end{corollary}
